---
title: "Publication profiles of Humanists"
format: pdf
editor: visual
---

```{r}
library(archetypes)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(geomtextpath)

```

# Description

```{r}
authors <- readr::read_delim("data/authors_metrics_dialnet.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)
```

Analysis of a set of 60,063 researchers from the Humanities. They are distributed among 40 different disciplines, but we should reshuffle them. Also there are NA values in some variables.

```{r}
# summary(authors)
authors <- authors %>%
  filter(!is.na(Age)) %>%
  filter(First_pub > 1949) %>%
  filter(Total_pubs>1) %>% # TEST
  filter(Total_pubs != max(Total_pubs)) #Remove the Pope

# Group discipline into broader fields
disciplina_map <- data.frame( # Lookup table to match disc and fields
  Disciplina = c(
    "Antropología Social",
    "Arqueología",
    "Expresión Gráfica Arquitectónica",
    "Expresión Gráfica en la Ingeniería",
    "Historia del Arte",
    "Pintura",
    "Estudios Árabes e Islámicos",
    "Estudios de Asía Oriental",
    "Estudios Hebreos y Arameos",
    "Filología Alemana",
    "Filología Catalana",
    "Filología Eslava",
    "Filología Francesa",
    "Filología Gallega y Portuguesa",
    "Filología Griega",
    "Filología Inglesa",
    "Filología Italiana",
    "Filología Latina",
    "Filología Románica",
    "Filología Vasca",
    "Filosofía",
    "Filosofía Moral",
    "Lógica y Filosofía de la Ciencia",
    "Geografía Humana",
    "Ciencias y Técnicas Historiográficas",
    "Historia Antigua",
    "Historia Contemporánea",
    "Historia de América",
    "Historia de la Ciencia",
    "Historia del Pensamiento y de los Movimientos ...",
    "Historia Medieval",
    "Historia Moderna",
    "Prehistoria",
    "Lengua Española",
    "Lingüística General",
    "Lingüística Indoeuropea",
    "Literatura Española",
    "Música",
    "Paleontología",
    "Traducción e Interpretación"
  ),
  field = c(
    "Anthropology",
    "Archaeology",
    "Arts",
    "Arts",
    "Arts",
    "Arts",
    "Cultural Studies",
    "Cultural Studies",
    "Cultural Studies",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philology",
    "Philosophy",
    "Philosophy",
    "Philosophy",
    "Geography",
    "History",
    "History",
    "History",
    "History",
    "History",
    "History",
    "History",
    "History",
    "History",
    "Language & Linguistics",
    "Language & Linguistics",
    "Language & Linguistics",
    "Literature",
    "Music",
    "Paleontology",
    "Translation Studies"
    )
  )

# Use lookup table to update authors$field
authors$field <- disciplina_map$field[match(authors$Disciplina, disciplina_map$Disciplina)]
```

Distribution by field.

```{r}
authors %>%
  group_by(field) %>%
  summarise(count = n(), age = mean(Age)) %>%
  ggplot(aes(x = reorder(field, count), y = count, fill = age)) +
  geom_bar(stat = "identity") +
  labs(x = "Field", y = "No. academics") +
  theme_classic() +
  coord_flip()
```

Evolution per age.

```{r}
ggplot(authors, aes(x = First_pub)) +
       geom_histogram(color = "black", fill = "darkgrey") +
  xlim(1950, 2020) +
  labs(x ="First publication year", y = "No. academics") +
  theme_classic()
```

One would expect age to be highly correlated with number of publications.

```{r}
cor(authors$Age, authors$Total_pubs)
```

```{r}
plot(authors$Age, authors$Total_pubs)
```

Publication profile by age and by field

# Archetypes

Create distribution.

```{r}
# Description
authors %>%
  select(
    `CODIGO AUTOR DIALNET`,
    Article,
    Book,
    Chapter,
    `Proceeding paper`,
    `Edited-proceedings`,
    `Edited-book`,
    Total_pubs,
    Inter,
    Local
  ) %>%
  tidyr::pivot_longer(
    cols = !`CODIGO AUTOR DIALNET`,
    names_to = "variable",
    values_to = "value"
  ) %>%
  group_by(variable) %>%
  summarise(n = n(), mean(value), sd = sd(value))
```

I am going to define the variables and conduct the archetypal analysis based on two dimensions:

-   **Type of outputs.** That is, the publication types in which they communicate their findings. we consider 6 publication types:

    -   Books

    -   Book chapters

    -   Proceedings papers

    -   Edited books

    -   Edited proceedings

-   **Outreach.** That is, the international scope of their outputs. We consider two variables:

    -   Publications indexed in WoS and Scopus

    -   Non-English publications

```{r}
arcdf.2dim <- authors %>%
  select(Article, Book, Chapter, `Proceeding paper`, `Edited-proceedings`, `Edited-book`, Inter, Local, Total_pubs)

names(arcdf.2dim) <- c("Article", "Book", "Chapter", "Proceedingpaper", "Editedproceedings", "Editedbook", "Inter", "Local", "Total_pubs")

# Create proportions of everything

arcdf.2dim$papers <- arcdf.2dim$Article/arcdf.2dim$Total_pubs
arcdf.2dim$books <- arcdf.2dim$Book/arcdf.2dim$Total_pubs
arcdf.2dim$chapters <- arcdf.2dim$Chapter/arcdf.2dim$Total_pubs
arcdf.2dim$proceedings <- arcdf.2dim$Proceedingpaper/arcdf.2dim$Total_pubs
arcdf.2dim$editor <- arcdf.2dim$Editedbook / arcdf.2dim$Total_pubs
arcdf.2dim$confeditor <- arcdf.2dim$Editedproceedings / arcdf.2dim$Total_pubs
arcdf.2dim$mainstream <- arcdf.2dim$Inter/arcdf.2dim$Total_pubs
arcdf.2dim$local <- arcdf.2dim$Local/arcdf.2dim$Total_pubs

arcdf.2dim <- arcdf.2dim %>%
  select(papers, books, chapters, proceedings, editor, confeditor, mainstream, local)
```

Now let's run archetypal analysis and look at the ideal number of archetypes. Based on the RSS and following an elbow rule, seems to be 3.

```{r}
set.seed(1986)
arcfull1 <- stepArchetypes(
  data = arcdf.2dim, k =1:5, verbose = T, nrep = 3)

# See ideal number of archetypes based on RSS
screeplot(arcfull1)
```

Let's look at the parameters for three archetypes:

```{r}
best_model <- 3
arcfull3 <- bestModel(arcfull1[[best_model]])

arcall.params <- t(parameters(arcfull3)) # parameters
arcall.alpha <- cbind.data.frame(authors$`CODIGO AUTOR DIALNET`, arcfull3[["alphas"]])
names(arcall.alpha) <- # Alpha scores
  c("dialnet_id", "archetype1", "archetype2", "archetype3")
arcall.alpha$field <- "All fields"

barplot(arcfull3, as.data.frame(arcdf.2dim), percentiles = F)
```
